Viscous Erosion in Porous Media

### 2018-02-26 docs/methods/erosion_methods.tex
NM

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.
On each body, $\nn$ and $\ss$ denote the unit normal and unit tangent vectors respectively. Throughout the paper, we adopt the convention that $\ss$ points in the counterclockwise direction and $\nn$ points out of the fluid domain (i.e.~into the bodies). Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_\ell(s,t) \in
\gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align*}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.
\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_\ell$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a Hagen-Poiseuille flow
    is imposed.  The bodies are eroded at a rate that is proportional to
    the magnitude of the shear stress applied by the Stokesian fluid.
    The bodies are constrained to the middle third of the channel that
    is located between the dashed lines.}
\end{figure}

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.9 \textwidth]{./figs/area_drag.pdf}
\caption{The area and drag of an eroding body: (a) The body area measured in the simulations (black) shown against the scaling-law prediction \eqref{area_predict} (dotted red). The two are nearly indistinguishable. (b) The horizontal drag (black) on the body as it erodes, along with the viscous (green) and pressure (blue) contributions. The scaling prediction can be rearranged to give estimate \eqref{viscdrag} for the viscous drag (dotted red).}
\label{area_drag}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.85 \textwidth]{./figs/arangle.pdf}
\caption{Shape characteristics of an eroding body: (a) The aspect ratio, defined as the width to height, versus time. (b) The measured opening angle versus time. For both measurements, we show three cases of the smoothing parameters: $\eps = \sigma = 20/1024$, $\eps = \sigma = 10/1024$, and $\eps = \sigma = 5/1024$. In the most resolved simulation (the case 5/1024) the aspect ratio converges to 2.7 and the opening angle converges to the predicted value of $102^{\circ}$.}
\label{fig:arangle}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024c5
% Uses 01circ1024.in with epsfac = 5, sigfac = 5, dt = 2e-4, fixarea = 1, fixpdrop = 0

>>>>

% OTHER TITLE POSSIBILITIES
% A fast and accurate method for viscous erosion in a 2D porous medium

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We only
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write $\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.
On each body, $\nn$ and $\ss$ denote the unit normal and unit tangent vectors respectively. Throughout the paper, we adopt the convention that $\ss$ points in the counterclockwise direction and $\nn$ points out of the fluid domain (i.e.~into the bodies). Neglecting inertial forces, the dynamics of the fluid is
fully characterized by the position of the bodies $\xx_\ell(s,t) \in
\gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we enforce the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align*}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

\begin{figure}[htpb]
  \centering
  \input{figs/schematic.tikz}
  \caption{\label{fig:schematic} A schematic for the governing
    equations.  A no-slip boundary condition is imposed on each body
    $\gamma_\ell$ whose unit normal points outward relative to the
    geometry.  On the outer geometry $\Gamma$, a Hagen-Poiseuille flow
    is imposed.  The bodies are eroded at a rate that is proportional to
    the magnitude of the shear stress applied by the Stokesian fluid.
    The bodies are constrained to the middle third of the channel that
    is located between the dashed lines.}
\end{figure}

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.85 \textwidth]{./figs/area_drag.pdf}
\caption{The area and drag of an eroding body: (a) The body area measured in the simulations (black) shown against the scaling-law prediction \eqref{area_predict} (dotted red). The two are nearly indistinguishable. (b) The horizontal drag (black) on the body as it erodes, along with the viscous (green) and pressure (blue) contributions. The scaling prediction can be rearranged to give estimate \eqref{viscdrag} for the viscous drag (dotted red).}
\label{area_drag}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.85 \textwidth]{./figs/arangle.pdf}
\caption{Shape characteristics of an eroding body: (a) The aspect ratio, defined as the width to height, versus time. (b) The measured opening angle versus time. For both measurements, we show three cases of the smoothing parameters: $\eps = \sigma = 20/1024, 10/1024,$ and 5/1024. In the most resolved simulation (5/1024) the aspect ratio converges to 2.7 and the opening angle converges to the predicted value of $102^{\circ}$ in long time.}
\label{fig:arangle}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024c5
% Uses 01circ1024.in with epsfac = 5, sigfac = 5, dt = 2e-4, fixarea = 1, fixpdrop = 0
